\documentclass[11pt]{article}

\usepackage{geometry}               	% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                	% ... or a4paper or a5paper or ...
%\geometry{landscape}                	% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    	% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{pgf}
%\usepackage{epstopdf}
\usepackage{url}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
\setcounter{secnumdepth}{7}
\setcounter{tocdepth}{7}


\def\Hyperjaxb2{Hyperjaxb2}			%uniform definition for Hyperjaxb2
\def\Genmapp{GenMAPP}			%uniform definition for GenMAPP
\def\Hibernate{Hibernate}				%uniform definition for Hibernate



\title{XMLPipeDB: A Reusable Tool Chain for Building Relational Databases from XML Sources}
\author{
John David N. Dionisio \and
Joey Barrett \and
Joe Boyle \and
Adam Carasso \and
David Hoffman \and
Babak Naffas \and
Jeffrey Nicholas \and
Roberto Ruiz \and
Scott Spicer \and
Kam D. Dahlquist\\
\\
Loyola Marymount University
}
%\date{}                                           % Activate to display a given date or no date

\begin{document}

\maketitle

\abstract{
We set out to create a method of moving data from large genetic databases into the format used by \Genmapp. We limited the scope of our work to Uniprot and GO databases. PostgreSQL was used as the intermediate database. We were determined to use as many existing, open source tools as possible to facilitate this. We used \Hyperjaxb2 to facilitate generation of database schemas, \Hibernate~mapping files and target-specific \Hibernate~class files.  \Hibernate~was used to provide the object to relational mapping. We succeeded in builing a chain of tools to go from an XSD to a \Genmapp~database file. We also succeeded in building many reusable components that are not specific to nor tied solely to bioinformatics, Uniprot, GO or \Genmapp. The reusable tools created to give capability to convert an XSD to a database schema and supporting \Hibernate~files; configure a system to use \Hibernate~for database access, preview XML data, import XML data, and run queries on data in a database. The first tool is called xsd2db. The later three tools are collectively called xmlpipedbutils.  Three more tools, uniprotdb, godb and genmapp builder, were required to meet our end goal - generation of a \Genmapp~database file. This tool set is unique in that there is no open source tool set that uses \Hyperjaxb2~and \Hibernate~to accomplish such tasks.
}




\newpage
\tableofcontents
\newpage

\include{introduction}

\include{background}


\include{design}

%Adam%
\include{xsd2db}

%Babak, David, Jeffrey%
\include{xmlpipedbutils}

%Jeffrey%
\include{clientconfiguration}

%David%
\include{xmlimport}

%Babak%
\include{adhocqueries}

%Joe, Joey, Roberto, Scott%
\include{biodatabases}

%Joe, Joey%
\include{uniprot}

%Roberto, Scott%
\include{geneontology}

%Joey%
\include{genmappbuilder}


\section{Conclusion}
Despite the pitfalls encountered, incredibly tight timeframes and demands of assimilating an entirely new domain of knowledge, the XmlPipeDb team has produced an initial release of outstanding quality and utility. In each area, we have identified additional work to be done, whether to add polish, tweak bugs or add new functionality. The real test will come, however, when the biological community begins to use the tool chain and reap the rewards of our effort. Our since hope is that it will meet their needs and allow them to make more interesting discoveries about the building blocks of life. Of course, there will be issues and we will rise to meet the challenge. There will also be additional databases to support and format changes. These, too, we will strive to accomodate. Of course, one of the great advantages of this tool chain being open source is that it is not left only to us to do this work, but to anyone with the vision and the will to make it happen.



\pagebreak
\bibliography{../xmlpipedb-refs}
\bibliographystyle{alpha}

\end{document}
